GMU MS SWE Handbook\\{\large A Handbook For the Masters of Software Engineering program at George Mason University

### 2024-10-15 ennv.tex
update

The output of a DNN is obtained by iteratively computing  the  values  of  neurons  in  each  layer.
The value of a neuron in the input layer is the input data. The value of a neuron in the hidden layers is computed by applying an \emph{affine transformation} to values of neurons in the previous layers, then followed by an \emph{activation function} such as the popular Rectified Linear Unit (ReLU) activation.

For this activation, the value of a hidden neuron \(y\) is 
$ReLU(w_1v_1 + \dots{} + w_nv_n + b)$, where \(b\) is the bias parameter of \(y\), \(w_i, \dots, w_n\) are the weights of \(y\), \(v_1,\dots,v_n\) are the neuron values of preceding layer, \(w_1v_1 + \dots + w_nv_n+b\) is the affine transformation, and \(ReLU(x) = \max(x,0)\) is the ReLU activation. The values of a neuron in the output layer is evaluated similarly but it may skip the activation function.
A ReLU activated neuron is said to be \emph{active} if its input value is greater than
zero and \emph{inactive} otherwise.

\paragraph{DNN Verification} Given a DNN \(N\) and a property $\phi$, the \emph{DNN verification problem} asks if $\phi$ is a valid property of $N$.
Typically, $\phi$ is a formula of the form $\phi_{in} \Rightarrow \phi_{out}$, where $\phi_{in}$ is a property over the inputs of $N$ and $\phi_{out}$ is a property over the outputs of $N$.
%This form of properties has been used to encode safety and security requirements of DNNs, e.g., safety specifications to avoid collision in unmanned aircraft~\cite{kochenderfer2012next} and \emph{adversarial robustness}~\cite{katz2017towards} properties desired by all DNNs, in which a small input perturbation does not cause major spikes in the DNN's outputs.
A DNN verifier attempts to find a \emph{counterexample} input to $N$ that satisfies $\phi_{in}$ but violates $\phi_{out}$.  If no such counterexample exists, $\phi$ is a valid property of $N$. Otherwise, $\phi$ is not valid and the counterexample can be used to retrain or debug the DNN~\cite{huang2017safety}.

%\begin{wrapfigure}{r}{0.3\textwidth}
%    \centering
%    \includegraphics[width=1\linewidth]{figure/dnn.pdf}
%    \caption{\label{fig:dnn} An FNN with ReLU.}
%  \end{wrapfigure}
\paragraph{Example} Fig.~\ref{fig:dnn} shows a simple DNN with two inputs $x_1,x_2$, two hidden neurons $x_3,x_4$, and one output $x_5$. The weights of a neuron are shown on its incoming edges , and the bias is shown above or below each neuron. The outputs of the hidden neurons  are computed the affine transformation and ReLU, e.g., $x_3 = ReLU(-0.5x_1+0.5x_2+1.0)$. The output neuron is computed with just the affine transformation, i.e., $x_5=-x_3+x_4-1$.

A valid property for this DNN is that the output is $x_5 \le 0$ for any inputs $x_1 \in [-1,1], x_2\in[-2,2]$. An invalid property for this network is that $x_5 > 0$ for those similar inputs.
A counterexample showing this property violation is $\{x_1=-1, x_2=2\}$, from which the network evaluates to $x_5=-3.5$. Such properties can capture \emph{safety requirements} (e.g., a rule in an  collision avoidance system in~\cite{kochenderfer2012next,katz2017reluplex} is ``if the intruder is distant and significantly slower than us, then we stay below a certain threshold'') or \emph{local robustness}~\cite{katz2017towards} conditions (a form of adversarial robustness stating that small perturbations of a given input all yield the same output).

@@

\section{Affine Transformation}
The affine transformation (AF) of a neuron is the sum of the products of the weights of the incoming edges and the values of the neurons in the previous layer, plus the bias of the neuron.
More specifically, the AF of a neuron \(y\) with weights \(w_1, \dots, w_n\) and bias \(b\) and the values of neurons in the previous layer \(v_1, \dots, v_n\) is \(w_1v_1 + \dots + w_nv_n + b\).

For example, the AF of a neuron \(y\) with weights \(-0.5, 0.5\) and bias \(1.0\) and the values of neurons in the previous layer \(x_1, x_2\) is \(-0.5x_1 + 0.5x_2 + 1.0\).

@@

\begin{itemize}
\item ReLU (Rectified Linear Unit): ReLU is the most popular activation function in DNNs. It returns 0 if the input is less than zero, and the input itself otherwise. It is often used in hidden layers.\\
    \begin{center}
        $ReLU(x) = \max(x,0)$
    \end{center}
\item Sigmoid: Sigmoid is a smooth function that maps any real value to the range (0,1). It is often used in the output layer of a binary classification problem.\\
    \begin{center}
        $Sigmoid(x) = \frac{1}{1+e^{-x}}$
    \end{center}
\item Tanh: Tanh is similar to the sigmoid function but maps any real value to the range (-1,1). It is often used in the output layer of a multi-class classification problem.\\
    \begin{center}
        $Tanh(x) = \frac{e^x-e^{-x}}{e^x+e^{-x}}$
    \end{center}
\item Softmax: Softmax is a generalization of the sigmoid function that maps any real value to the range (0,1) and ensures that the sum of the output values is 1. It is often used in the output layer of a multi-class classification problem.\\
    \begin{center}
    $Softmax(x)_i = \frac{e^{x_i}}{\sum_{j=1}^{n}e^{x_j}}$
    \end{center}
\end{itemize}

@@

@@

\chapter{Verification of Neural Networks}\label{sec:verification}

>>>>

The output of a DNN is obtained by iteratively computing  the  values  of  neurons  in  each  layer.
The value of a neuron in the input layer is the input data. The value of a neuron in the hidden layers is computed by applying an \emph{affine transformation} (\autoref{sec:affine}) to values of neurons in the previous layers, then followed by an \emph{activation function} (\autoref{sec:activation}) such as the popular Rectified Linear Unit (ReLU) activation.

@@

\section{Affine Transformation}\label{sec:affine}
The affine transformation (AF) of a neuron is the sum of the products of the weights of the incoming edges and the values of the neurons in the previous layer, plus the bias of the neuron.
More specifically, the AF of a neuron \(y\) with weights \(w_1, \dots, w_n\) and bias \(b\) and the values of neurons in the previous layer \(v_1, \dots, v_n\) is \(w_1v_1 + \dots + w_nv_n + b\).

%AF is often applied to hidden layers of a DNN to compute the value of a neuron in the hidden layer.
%The values of a neuron in the output layer is evaluated similarly but it may skip the activation function.

For example, the AF of a neuron \(x_3\) in \autoref{fig:dnn} with (incoming arrows) weights \(-0.5, 0.5\) and bias \(1.0\) and the values of neurons in the previous layer \(x_1, x_2\) is \(-0.5x_1 + 0.5x_2 + 1.0\).

@@

\begin{itemize}
\item ReLU (Rectified Linear Unit): ReLU is the most popular activation function in DNNs. It returns 0 if the input is less than zero, and the input itself otherwise. It is often used in hidden layers and skipped in the output layer.  
A ReLU activated neuron is said to be \emph{active} if its input value is greater than
zero and \emph{inactive} otherwise.\\
    \begin{center}
        $ReLU(x) = \max(x,0)$
    \end{center}

\item Sigmoid: Sigmoid is a smooth function that maps any real value to the range (0,1). It is often used in the output layer of a binary classification problem.\\
    \begin{center}
        $Sigmoid(x) = \frac{1}{1+e^{-x}}$
    \end{center}
\item Tanh: Tanh is similar to the sigmoid function but maps any real value to the range (-1,1). It is often used in the output layer of a multi-class classification problem.\\
    \begin{center}
        $Tanh(x) = \frac{e^x-e^{-x}}{e^x+e^{-x}}$
    \end{center}
\item Softmax: Softmax is a generalization of the sigmoid function that maps any real value to the range (0,1) and ensures that the sum of the output values is 1. It is often used in the output layer of a multi-class classification problem.\\
    \begin{center}
    $Softmax(x)_i = \frac{e^{x_i}}{\sum_{j=1}^{n}e^{x_j}}$
    \end{center}
\end{itemize}

@@

\section{Example} 
Fig.~\ref{fig:dnn} shows a simple DNN with two inputs $x_1,x_2$, two hidden neurons $x_3,x_4$, and one output $x_5$. The weights of a neuron are shown on its incoming edges , and the bias is shown above or below each neuron. The outputs of the hidden neurons  are computed the affine transformation and ReLU, e.g., $x_3 = ReLU(-0.5x_1+0.5x_2+1.0)$. The output neuron is computed with just the affine transformation, i.e., $x_5=-x_3+x_4-1$.

\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{figure/dnn.pdf}
    \caption{\label{fig:dnn} An FNN with ReLU.}
\end{figure}

@@

\chapter{Verification of Neural Networks}\label{sec:verification}
\paragraph{DNN Verification} Given a DNN \(N\) and a property $\phi$, the \emph{DNN verification problem} asks if $\phi$ is a valid property of $N$.
Typically, $\phi$ is a formula of the form $\phi_{in} \Rightarrow \phi_{out}$, where $\phi_{in}$ is a property over the inputs of $N$ and $\phi_{out}$ is a property over the outputs of $N$.
%This form of properties has been used to encode safety and security requirements of DNNs, e.g., safety specifications to avoid collision in unmanned aircraft~\cite{kochenderfer2012next} and \emph{adversarial robustness}~\cite{katz2017towards} properties desired by all DNNs, in which a small input perturbation does not cause major spikes in the DNN's outputs.
A DNN verifier attempts to find a \emph{counterexample} input to $N$ that satisfies $\phi_{in}$ but violates $\phi_{out}$.  If no such counterexample exists, $\phi$ is a valid property of $N$. Otherwise, $\phi$ is not valid and the counterexample can be used to retrain or debug the DNN~\cite{huang2017safety}.

\paragraph{Example} A valid property for the DNN in \autoref{fig:dnn} is that the output is $x_5 \le 0$ for any inputs $x_1 \in [-1,1], x_2\in[-2,2]$. An invalid property for this network is that $x_5 > 0$ for those similar inputs.
A counterexample showing this property violation is $\{x_1=-1, x_2=2\}$, from which the network evaluates to $x_5=-3.5$. Such properties can capture \emph{safety requirements} (e.g., a rule in an  collision avoidance system in~\cite{kochenderfer2012next,katz2017reluplex} is ``if the intruder is distant and significantly slower than us, then we stay below a certain threshold'') or \emph{local robustness}~\cite{katz2017towards} conditions (a form of adversarial robustness stating that small perturbations of a given input all yield the same output).
